Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-29 manuscript/HIAs.qmd
calculate distance NF to Målselv

(positionMap <- tm_shape(outline) +
  tm_polygons() +
  tm_shape(centroids) +
  tm_text(
    text = "Municipality",
    size = 1,
    xmod = 2,
    ymod = -1
  ) +
  tm_symbols(
    size = 2,
    shape = 17,
    col = "black"
  ) +
  tm_grid(projection = 4326) +
  tm_layout(
    bg.color = "skyblue",
    outer.margins = c(0.01, .02, .02, .02)))
```

@@

<!--# add mean_HIA to the mix -->

@@

# Take muni3 and add all sorts of other data to it
# using left_join.
# Save local cached file
muni_tbl <- muni3 |>
  # calculate total area
  mutate(
    area_km =
      round(
        units::drop_units(
          geom |> st_area() * 1e-6
        )
      )
  ) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(terrestrial |> select(kommunenummer, t_area_km),
    keep = F
  ) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km / t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_nf_terra_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2 / mire_km2 * 100, 2)) |>
  left_join(infraMuni3_summary)

@@

We chose three municipalities in Norway to test out the indicators (@fig-studyLocation).
These municipalities are spread over xxx km north to south, and also differ in the amount of mire area, the total area surveyed, and the prevalence of infrastructure (@tbl-variables).

>>>>

(positionMap <- tm_shape(outline) +
  tm_polygons() +
  tm_shape(centroids) +
  tm_text(
    text = "Municipality",
    size = .8,
    xmod = 2,
    ymod = -1
  ) +
  tm_symbols(
    size = 1.5,
    shape = 17,
    col = "black"
  ) +
  tm_grid(projection = 4326) +
  tm_layout(
    bg.color = "skyblue",
    outer.margins = c(0.01, .02, .02, .02)))
```

```{r distanceBetweenMunis}
#| eval: true

# what is the distance between Nordre Follo and Målselv
(latlong <- centroids |>
  st_transform(4326) |>
st_coordinates() |>
  as_tibble())

round(max(latlong$Y)-min(latlong$Y), 1)

```

@@

@@

# Take muni3 and add all sorts of other data to it
# using left_join.
# Save local cached file
muni_tbl <- muni3 |>
  # calculate total area
  mutate(
    area_km =
      round(
        units::drop_units(
          geom |> st_area() * 1e-6
        )
      )
  ) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(terrestrial |>
              as_tibble() |> 
              select(kommunenummer, t_area_km),
    keep = F
  ) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km / t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_nf_terra_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2 / mire_km2 * 100, 2)) |>
  left_join(infraMuni3_summary)

@@

We chose three municipalities in Norway to test out the indicators (@fig-studyLocation).
These municipalities are spread over `r round(max(latlong$Y)-min(latlong$Y), 1)` degrees latitude, and also differ in the amount of mire area, the total area surveyed, and the prevalence of infrastructure (@tbl-variables).
